---
title: "tom-index"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{tom-index}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup, include=FALSE, echo = FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(plotly)
library(crosstalk)
```

## Background

For this project, I decided to recreate Hans Rosling's population and GDP graph. The graph can be found [here](https://www.youtube.com/watch?v=jbkSRLYSojo&ab_channel=BBC). 

Since the video did not share the dataset, I downloaded another dataset from [Our World in Data](https://ourworldindata.org/grapher/life-expectancy-vs-gdp-per-capita). 

## Data Visualization

I used plotly and crosstalk to make a visualization.  I think crosstalk does not go well with animation plots as the continents/countries shown are different from ones selected.

```{r message=FALSE, warning=FALSE, echo = FALSE}
shared_le <- SharedData$new(le)

bscols(widths = c(2,NA),
  list(
    filter_select("continent", "Continent", shared_le, ~continent)
  ),
  ggplotly(shared_le %>% 
  ggplot(aes(gdp_per_capita, life_expectancy)) +
  geom_point(aes(size = population_historical_estimates, frame = year, ids = entity, color = continent)) +
  scale_x_log10() +
  labs(x = "GDP per Capita", y = "Life Expectancy", size = ""))
)
```

## Show countries in North America

```{r message=FALSE, warning=FALSE, echo = FALSE}
le_na <- le %>% filter(continent == "North America")
shared_le_na <- SharedData$new(le_na)

bscols(widths = c(2,NA),
  list(
    filter_select("entity", "Country", shared_le_na, ~entity)
  ),
  ggplotly(
  ggplot(data = shared_le_na, aes(gdp_per_capita, life_expectancy)) +
  geom_point(aes(size = population_historical_estimates, frame = year, color = entity)) +
  scale_x_log10() +
  labs(x = "GDP per Capita", y = "Life Expectancy", color = "Country", size = ""))
)
```

## Better One

```{r message=FALSE, warning=FALSE, echo = FALSE}
bscols(widths = c(2,4,6),
  list(
    filter_select("entity", "Country", shared_le, ~entity),
    filter_slider("year", "Years", shared_le, ~year, round = TRUE)
  ),
  add_lines(shared_le %>%
  plot_ly(x = ~year, y = ~life_expectancy, color = ~entity, id = ~entity), showlegend = FALSE) %>% 
    layout(xaxis = list(title = "Year"), yaxis = list(title = "Life Expectancy")),
  
  add_lines(shared_le %>%
  plot_ly(x = ~year, y = ~gdp_per_capita, color = ~entity, id = ~entity) %>% 
    layout(xaxis = list(title = "Year"), yaxis = list(title = "GDP Per Capita")))
)
```

## Improvement

For the improvement, I would like to make a similar graph to the Figure 13.10 from Wilke's [Fundamentals of Data Visualization](https://clauswilke.com/dataviz/time-series.html).

The code was taken and modified from [Wilke's github repo](https://github.com/clauswilke/dataviz/blob/master/time_series.Rmd).

The graph below shows how life expectancy and GDP per capita transition over time in Afghanistan. The transparency of the line represents year. In the future, I would like to incorporate the graph with crosstalk. It seems like crosstalk does not go well with `geom_text_repel`. I also need to figure out how to show only selected country in the graph (in the previous graphs, `color = entity` does it for me, but the graph below has nothing to differentiate countries if that makes sense).

```{r message=FALSE, warning=FALSE, echo = FALSE}
le2 <- le %>% 
  mutate(label = case_when
         (year == "1950" ~ "1950",
          year == "1960" ~ "1960",
          year == "1970" ~ "1970",
          year == "1980" ~ "1980",
          year == "1990" ~ "1990",
          year == "2000" ~ "2000",
          year == "2010" ~ "2010",
          year == "2018" ~ "2018"))
```

```{r message=FALSE, warning=FALSE, echo = FALSE}
library(ggridges)
library(lubridate)
library(ggrepel)

time_series <- le2 %>%
  filter(entity == "Afghanistan") %>% 
  ggplot() +
    aes(gdp_per_capita, life_expectancy, colour = as.numeric(year)) + 
    geom_path(size = 1, lineend = "round") +
  geom_text_repel(
    aes(label = label), point.padding = .2, color = "black",
    min.segment.length = 0, size = 12/.pt,
    nudge_x = le2$nudge_x,
    nudge_y = le2$nudge_y,
    direction = "y"
  ) + 
  scale_colour_gradient(low = "#E7F0FF", high = "#035B8F") + #"#0072b2") +
  guides(colour = FALSE) +
  coord_cartesian(clip = "off") +
  theme(
    axis.ticks.length = unit(0, "pt"),
    plot.margin = margin(21, 14, 3.5, 1.5)) +
  labs(title = "Life Expectancy and GDP change over time in Afghanistan", y = "Life Expectancy", x = "GDP per Capita") +
  theme_bw()

time_series
```

## Reflection

* I wanted to make it as simple as possible. Without the selector for continents/countries, all dots were crammed together and it was difficult to see specific continents/countries. After implementing the selector, I think the visualization became much simpler.

* For the animation visualizations, selected countries do not match with what are shown in the graph, which I want to fix in the future. For the bar graph, the adjacent countries in the list have similar color, so it is hard to distinguish.

* The most frustrating part was definitely the mismatching between selected and shown.

